---
title: "Sapling abundance and transition rates"
format: gfm
editor: source
---

```{r}
library(duckdb)
library(dplyr)
library(ggplot2)

source(here::here("R", "query_tables_db_fxns.R"))
source(here::here("use_cases", "ingrowth", "sapling_functions.R"))

con <- connect_to_tables(here::here("data", "db", "forestTIME-cli.duckdb"))

```

# Sapling presence

We can pull saplings for a state (or any other set of filter variables) with this function:

```{r}

ct_saplings <- generate_sapling_table(con = con,
                                      tree_id_method = "composite",
                                      conditions = create_conditions(STATECD == 9))

knitr::kable(head(ct_saplings))

```

This table contains the necessary information to compute the abundance or biomass of saplings on a (sub)plot, and includes the `TPA_UNADJ` column for scaling to the whole plot scale. 

# Transition rates

We can pull a sapling transition table for Connecticut:

```{r}

ct_transitions <- generate_sapling_transition_table(con = con,
                                                    tree_id_method = "composite",
                                                    conditions = create_conditions(STATECD == 9))

knitr::kable(head(ct_transitions))
```

And look at the distribution of death and recruitment rates:

```{r}

ggplot(ct_transitions, aes(presumed_dead_nb_prop)) +
  geom_histogram(boundary = 0)

ggplot(ct_transitions, aes(sapling_tree_nb_prop)) +
  geom_histogram(boundary = 0)

```

You can also pull for, for example, just one species:

```{r}
mn_red_maple_transitions <- generate_sapling_transition_table(con = con,
                                                    tree_id_method = "composite",
                                                    conditions = create_conditions(STATECD == 27,
                                                                                   SPCD == 316,
                                                                                   SPCDS == 1)) # this removes some records of trees changing species, which mess up the proportion sums

knitr::kable(head(mn_red_maple_transitions))


```

```{r}
dbDisconnect(con, shutdown = TRUE)

```

