---
title: "Comparison of daisy chain and tree number"
format: 
  gfm: 
    df-print: kable
editor: source
---

## Context

The "daisy chain" method is to link backwards in time using `PREV_TRE_CN` and `CN`. The "tree number" method is to concatenate columns `STATECD`, `UNITCD`, `COUNTYCD`, `PLOT`, `SUBP`, and `TREE` to get a unique tree number for each tree. In principle these _should_ give equivalent outcomes, but we don't know if they really always do!

## Code 

```{r}
#| label: setup

library(arrow)
library(dplyr)
library(ggplot2)

state_number <- c(4, 9, 27)
arrow_dir <- "data/arrow"

```


```{r}
#| label: create unique number table

tree_unique_number <- open_dataset(
  here::here(arrow_dir, "TREE_RAW"),
  partitioning = c("STATECD", "COUNTYCD"),
  format = "csv",
  hive_style = T,
  col_types = schema(
    CN = float64(),
    TREE_FIRST_CN = float64()
  )) |>
  filter(STATECD %in% state_number) |>
  mutate(TREE_UNIQUE_NUM = paste(STATECD,
                                 UNITCD,
                                 COUNTYCD,
                                 PLOT,
                                 SUBP,
                                 TREE, 
                                 sep = "_")) |>
  compute()
```


```{r}
#| label: open daisy chain table

join_cns <-  open_dataset(
  here::here(arrow_dir, "TREE_CN_JOIN"),
  partitioning = c("STATECD", "COUNTYCD"),
  format = "csv",
  hive_style = T,
  col_types = schema(
    CN = float64(),
    TREE_FIRST_CN = float64()
  )) |>
  filter(STATECD %in% state_number) |>
  compute()

```


```{r}
#| label: create unmatched_cns table

unmatched_cns <- join_cns |>
  left_join(tree_unique_number) |>
  select(TREE_FIRST_CN, TREE_UNIQUE_NUM) |> 
  distinct() |>
  collect() |> 
  group_by(TREE_FIRST_CN) |>
  mutate(n_NUM = n()) |>
  group_by(TREE_UNIQUE_NUM) |>
  mutate(n_FIRST_CN = n()) |>
  ungroup() |>
  filter(n_NUM > 1 |
           n_FIRST_CN > 1)

unmatched_deeper_dives <- join_cns |>
  left_join(tree_unique_number) |>
  filter(TREE_UNIQUE_NUM %in% unmatched_cns$TREE_UNIQUE_NUM) |>
  arrange(TREE_UNIQUE_NUM, CYCLE) |>
  collect() |>
  mutate(TREE_FIRST_CN = as.character(TREE_FIRST_CN),
         CN = as.character(CN),
         PREV_TRE_CN = as.character(PREV_TRE_CN)) |>
  group_by(TREE_UNIQUE_NUM) |>
  mutate(CN_break = TREE_FIRST_CN != CN[1]) |>
  mutate(prev_break = lag(CN_break),
         prev_status = lag(STATUSCD)) |>
  mutate(is_break_point = ifelse(CN_break, ifelse(prev_break, FALSE, TRUE), FALSE)) 


```

There are `r nrow(unmatched_cns)` instances where there isn't a perfect 1-1 matching of one TREE_UNIQUE_NUM to TREE_FIRST_CN. 
For these, it's always that 1 TREE_UNIQUE_NUM has matched to multiple TREE_FIRST_CNs. 

## Mismatches

### CT

CT has no mismatches.

### MN

In MN, each of these instances occurred when a tree had a `STATUSCD = 0` in one cycle and then received a new `CN` and a `PREV_TRE_CN = NA` in the following cycle. 
That is, in the following cycle, the previous tree CN was forgotten and the chain was broken. 

`STATUSCD = 0` means a tree is not part of the current sample, e.g. due to being incorrectly tallied or on a dangerous or inaccessible condition. 

In MN, _all_ of the breaks occurred on plot 21085, subplots 1 and 2, in 2012. Perhaps something was the matter with those plots in 2012?

Below is tables showing this.

```{r}

mn_deeper_dives <- unmatched_deeper_dives |>
  filter(STATECD == 27)

knitr::kable(mn_deeper_dives)

knitr::kable(mn_deeper_dives |>
  filter(is_break_point)
)

```


### AZ

The Arizona mismatches (approx. 1600 trees, of 71000 total) do not follow the same pattern as MN. 
All but 5 of these instances occur in CYCLE 3 between 2001-2005 when there are 2 visits to a tree within the same cycle. 
For those trees, the second visit to the tree has `PREV_TRE_CN = NA`, so there is no link to the first visit. 
To my understanding, the sampling methodology only calls for one visit per tree per cycle. 

Of the remaining 5 instances, I have no explanation.
But, I'll note that in 4 of those 5, the second record for a tree is of a different species than the first, suggesting some error somewhere. 

```{r}

#| arizona

az_deeper_dives <- unmatched_deeper_dives |>
  filter(STATECD == 4)

az_repeat_visits <- tree_unique_number |>
  filter(STATECD == 4) |>
  group_by(CYCLE, TREE_UNIQUE_NUM) |>
  arrange(INVYR) |>
  mutate(NVISITS = n(),
         VISIT_NUMBER = row_number(),
         CN = as.character(CN)) |>
  filter(NVISITS > 1,
         VISIT_NUMBER == 1) |>
  collect()

all(az_repeat_visits$CN %in% az_deeper_dives$CN)

no_repeats <- az_deeper_dives |> 
  anti_join(az_repeat_visits, by = "CN") |>
  group_by(TREE_UNIQUE_NUM) |>
  mutate(N_CNS = length(unique(TREE_FIRST_CN))) |>
  filter(N_CNS > 1)

knitr::kable(no_repeats)

```
