---
title: "Extracting FIA timeseries"
format: gfm
editor: source
---

```{r}
#| echo: FALSE
#| message: FALSE
#| warning: FALSE

library(arrow)
library(dplyr)
library(ggplot2)

source(here::here("R", "demo_functions.R"))

```

## 1. `select_trees`

`select_trees()` extracts persistent IDs for all trees that meet user-supplied criteria.

In this example, the criteria include:

- `STATECD`
- `COUNTYD`
- `PLOT`
- `SPCD`
- `min_years`, the minimum number of years between the first and last measurement
- `min_measurements`, the minimum number of survey visits to that tree 

::: {.callout-note}

What additional filtering criteria should there be?

:::

By default, `select_trees` returns all trees:

```{r}

my_trees <- select_trees()

nrow(my_trees)

knitr::kable(head(my_trees))

```

Passing additional options filters the trees:

```{r}

# Extract all red maples (SPCD = 316)

my_red_maples <- select_trees(spcd = 316)

nrow(my_red_maples)

knitr::kable(head(my_red_maples))

```

```{r}

# Extract red maples on plot 20010

my_red_maples_one_plot <- select_trees(spcd = 316,
                                   plot = 20010)

nrow(my_red_maples_one_plot)

knitr::kable(head(my_red_maples_one_plot))

```


## `get_timeseries`

`get_timeseries` extracts measurements for a selection of trees. 
By default, it will return all measurements of `DIA`, `HT`, and `STATUSCD` for the given trees.
The time window included can be modified using the `min_year` and `max_year` arguments. 
The measurements returned can be modified to add additional variables using the `variables` argument.
Currently, any of the columns of the `TREE` table can be added as additional variables.

::: {.callout-note}

What should be the default measurements provided?

What additional options should there be, including options from other tables?

:::


For example, extracting the default measurements for all years for red maples on plot 20010:

```{r}

my_maple_timeseries <- get_timeseries(my_red_maples_one_plot)

nrow(my_maple_timeseries)

knitr::kable(head(my_maple_timeseries))
```



Modifying the time window and variables to return:


```{r}

my_maple_timeseries2 <- get_timeseries(my_red_maples_one_plot,
                                            min_year = 2010,
                                            max_year = 2023,
                                            variables = c("DIA", "HT", "ACTUALHT"))

nrow(my_maple_timeseries2)

knitr::kable(head(my_maple_timeseries2))
```

Here is a plot of how individual maple trees' diameters have changed over time:

```{r}
#| label: fig-mapledia
#| fig-cap: Diameters of red maple trees on plot 20010 over time. Individual lines represent individual trees.
#| echo: false

ggplot(my_maple_timeseries, aes(INVYR, DIA, group = as.factor(TREE_FIRST_CN))) +
  geom_line() +
  geom_point() +
  theme_bw() +
  ggtitle("Maple tree diameters over time", subtitle = "Plot 20010") +
  xlab("INVYR") +
  ylab("DIA")

```