---
title: "Carbon estimation"
format: gfm
editor: source
---

```{r}
library(ggplot2)

source(here::here("R", "query_tables_db_fxns.R"))
```


# Getting some data

This code pulls 6 trees out of Idaho to start working with, and adds a row to match the Douglas fir example worked in the text.

Note that I am pretending Idaho is in DIVISION 240, which is wrong, to match the worked example.

```{r, eval = F}
con <- connect_to_tables(here::here("data", "db", "foresttime-from-state-parquet.duckdb"))

some_trees <- query_tree_surveys(con,
                                 conditions = create_conditions(STATECD == 16,
                                                                !is.na(DIA),
                                                                !is.na(HT),
                                                                STATUSCD == 1,
                                                                !is.na(SPCD),
                                                                SPCD %in% c(202, 231, 73)),
                                 variables = c("STANDING_DEAD_CD", "DECAYCD", "DIA", "HT")) |>
  group_by(SPCD) |>
  mutate(rowindex = row_number()) |>
  ungroup() |>
  filter(rowindex <= 2) |>
  select(-rowindex) |>
  head() |>
  bind_rows(data.frame(SPCD = 202,
                       HT = 110,
                       DIA = 20))

dbDisconnect(con, shutdown = TRUE)

write.csv(some_trees, here::here("use_cases", "carbon_estimation", "toy_tree_data.csv"), row.names = FALSE)

```

```{r}

some_trees <- read.csv(here::here("use_cases", "carbon_estimation", "toy_tree_data.csv"))

```

# Step 1 Gross total stem wood volume

## Identify the model for gross total stem wood volume

Identify the right model and coefficients to use based on species code and division, using Table S1A from the supplement ([download here](https://www.fs.usda.gov/research/publications/gtr/gtr_wo104/WO-GTR-104-Supp1.zip)).


```{r}

table_s1a <- read.csv(here::here("use_cases", 
                                 "carbon_estimation",
                                 "Supplemental_tables",
                                 "Table S1a_volib_coefs_spcd.csv"))

```

First we need to get division. As far as I can tell this is only available from a map - is there a data table for this?

Also note that not all species have different models across divisions. Ideally we would create a table that has SPCD, STATE (or PLOT), and DIVISION. DIVISION should be "" if it doesn't matter. This will be a little work to create. 

And, not all species are represented in Table S1a. From this sample, we lose species 231. From the paper:

"For species not included in the spcd tables, the jenkins suffix tables are used with model 5 and associated coefficients for the Jenkins group that represents the species of interest.
Species assignments to Jenkins groups are in FIADB REF_SPECIES table as variable name JENKINS_SPGRPCD. 
Note that Jenkins group coefficients incorporate the predicted random effect into the reported coefficients, i.e., in some cases, the value is a sum of the fixed and random effects."

Here I get DIVISION-SPCD-model information for the 3 species that happen to be in this toy dataset. 

```{r}

table_s1a_models <- table_s1a |>
  filter(SPCD %in% some_trees$SPCD,
         DIVISION %in% c("", "240")) |>
  # This chunk preferentially pulls the model specific to this division
  # if division is provided at all for each species.
  mutate(DIVISION_nchar = nchar(DIVISION)) |>
  group_by(SPCD) |>
  mutate(longest_division = max(DIVISION_nchar)) |>
  ungroup() |>
  filter(DIVISION_nchar == longest_division) |>
  select(-DIVISION_nchar, -longest_division)

```

And then re-attach it to the tree data dataset:

```{r}

some_trees_models <- left_join(some_trees, table_s1a_models)

```

### Notes 

Ideally for this step we would have a table of SPCD | PLOT | MODEL | coefficients. 

To create this we would need:

- a data table of which plots (or other locations) are in which divisions - it's not feasible to get this by manually examining the map
- the Jenkins codes table
- some code to join these together appropriately

This would be a great thing to see if Grant and Brian already have.


## Apply the model for gross total stem wood volume

This table tells us which of the 6 models described in the paper to use for this species in this location. It makes sense to me to code each model as a function. Here are functions for the models:

```{r}

model1 <- function(DIA, HT, a, b, c) {
  
  a * (DIA ^ b) * (HT^c)
  
}

model2 <- function(DIA, HT, SPCD, a, b, b1, c) {
  
  if(SPCD <300) {
    k <- 9
  } else {
    k <- 11
  }
  
  if(DIA < k) {
    
    a * (DIA^b) * (HT^c)
    
  } else {
    
    a * (k^(b - b1)) * (DIA^b1) * (HT^c)
    
  }
  
}

model3 <- function(DIA, HT, a, a1, b, c, c1) {
  
  a * (DIA ^ (a1 * ((1 - exp(-b*DIA)) ^ c1))) * (HT^c)
  
}

model4 <- function(DIA, HT, a, b, b1, c) {
  
  a * (DIA ^ b) * (HT^c) * exp(-b1 * DIA)
  
}

model5 <- function(DIA, HT, WDSG, a, b, c){
  
  a * (DIA^ b) * (HT^c) * WDSG
  
}

```

Note also that model5 requires species-level wood specific gravity (WDSG).

```{r}


some_trees_gtswv <- some_trees_models |>
  group_by_all() |>
  mutate(gtswv = ifelse(model == 1,
                        model1(DIA = DIA,
                               HT = HT,
                               a = a, 
                               b = b,
                               c = c),
                        ifelse(model == 2,
                           model2(DIA = DIA,
                                  HT = HT,
                                  SPCD = SPCD,
                                  a = a,
                                  b = b,
                                  b1 = b1,
                                  c = c),
                           NA)))
```

Run the math, and we get the right result for the Doug fir example:

```{r}

some_trees_gtswv <- some_trees_models |>
  group_by_all() |>
  mutate(gtswv = ifelse(model == 1,
                        model1(DIA = DIA,
                               HT = HT,
                               a = a, 
                               b = b,
                               c = c),
                        ifelse(model == 2,
                           model2(DIA = DIA,
                                  HT = HT,
                                  SPCD = SPCD,
                                  a = a,
                                  b = b,
                                  b1 = b1,
                                  c = c),
                           NA)))

knitr::kable(some_trees_gtswv)
```

# Step 2 Gross total stem bark volume

We get the models and coefficients from table S2a in the Supplement:

```{r}

table_s2a <-  read.csv(here::here("use_cases", 
                                 "carbon_estimation",
                                 "Supplemental_tables",
                                 "Table S2a_volbk_coefs_spcd.csv"))
```

```{r}
table_s2a_models <- table_s2a |>
  filter(SPCD %in% some_trees$SPCD,
         DIVISION %in% c("", "240")) |>
  mutate(DIVISION_nchar = nchar(DIVISION)) |>
  group_by(SPCD) |>
  mutate(longest_division = max(DIVISION_nchar)) |>
  ungroup() |>
  filter(DIVISION_nchar == longest_division) |>
  select(-DIVISION_nchar, -longest_division)

gtsbv_models <- left_join(some_trees, table_s2a_models)

some_trees_gtsbv <- gtsbv_models |>
  group_by_all() |>
  mutate(gtsbv = ifelse(model == 1,
                        model1(DIA = DIA,
                               HT = HT,
                               a = a, 
                               b = b,
                               c = c),
                        ifelse(model == 2,
                           model2(DIA = DIA,
                                  HT = HT,
                                  SPCD = SPCD,
                                  a = a,
                                  b = b,
                                  b1 = b1,
                                  c = c),
                           NA)))

knitr::kable(some_trees_gtsbv)

```

Note that the estimate for the Doug fir example diverges subtly from the paper (which has 13.191436232306). I think this is because the coefficients in the paper for SPCD 202/DIVISION 240 differ from those in table s2a. Specifically, `a` is 0.000031886237 in the paper and 0.000031900 in the table - truncation?

```{r}

model1(20, 110, a = 0.000031886237, b = gtsbv_models$b[7], c = gtsbv_models$c[7])

```

# Step 3 Gross total stem outside-bark volume

This is just a sum:

```{r}

some_trees_gtsobv <- some_trees |>
  mutate(gtsobv = some_trees_gtsbv$gtsbv +
                      some_trees_gtswv$gtswv)

knitr::kable(some_trees_gtsobv)

```

# Step 4 Heights to merchantable top

```{r}



```