---
title: "Comparison of daisy chain and tree number"
format: 
  gfm: 
    df-print: kable
editor: source
execute:
  echo: false
  output: true
---

## Context

The "daisy chain" method is to link backwards in time using `PREV_TRE_CN` and `CN`. The "tree number" method is to concatenate columns `STATECD`, `UNITCD`, `COUNTYCD`, `PLOT`, `SUBP`, and `TREE` to get a unique tree number for each tree. In principle these _should_ give equivalent outcomes, but we don't know if they really always do!

## States and trees affected 

```{r, include = F, output =F}
#| label: setup

library(arrow)
library(dplyr)
library(ggplot2)

state_situations <- read.csv(here::here("reports", "state_situations.csv")) |>
  mutate(STATEFP = as.character(STATEFP)) |>
  mutate(STATEFP = ifelse(nchar(STATEFP) == 1, paste0("0", STATEFP), STATEFP)) |>
  select(-STATENS) |>
  mutate(proportion_affected = round(100 * n / n_trees_overall, digits = 2))



states_map <- tigris::states() |>
  left_join(state_situations) |>
  filter(STATE != "HI",
         STATE != "AK")

theme_set(theme_void())

```

### How many states are affected?

```{r}

knitr::kable(state_situations |> 
  mutate(is_affected = n > 0) |>
  group_by(is_affected) |>
  tally())

```

### How many trees/state are affected?

```{r}

ggplot(states_map) + 
  geom_sf(aes(fill = proportion_affected)) +
  coord_sf(lims_method = "geometry_bbox") +
  scale_fill_gradient(name = "% of trees affected", trans = "log") +
  theme(legend.position = "bottom", legend.direction = "vertical")

```


### What (generally) describes the affect?

```{r}


ggplot(states_map) + 
  geom_sf(aes(fill =description)) +
  coord_sf(lims_method = "geometry_bbox") +
  theme(legend.position = "bottom", legend.direction = "vertical")
```
