---
title: "Trees shrink"
format: gfm
editor: source
---

# Context

After the 4/17 meeting, I (Renata) was curious about instances where trees appear to shrink over time.
This could be "true", measurement error, or something else.
This document is me digging deeper into those records. 
Specifically, I checked:

* That shrinking is occurring in the right direction in time (i.e. the years aren't reversed or something). **This checked out**.
* Whether shrinking seems to be associated with trees dying. **About 1/3 of the times a tree shrank, it also was marked dead.**
* Whether shrinking seems to be associated with disturbance. **Another ~1/3 of the time, a tree had a damage or disturbance code when it shrank.**
* The magnitude of shrinking - small or large decreases in size? **Fairly qualitatively, I don't see an association between death/distrubance/damage and larger decreases in size. And there are some pretty big decreases in size not associated with any of those explanations.**


```{r}
library(ggplot2)

source(here::here("R", "query_tables_db_fxns.R"))
source(here::here("R", "query_annualized.R"))
```


# Pulling records

```{r}


con <- connect_to_tables(here::here("data", "db", "foresttime-to-share.duckdb"))

```

```{r}


az_nfs <- query_annualized(
  con,
  conditions = create_conditions(
    STATECD == 4,
    OWNCD == 11,
    MULTI_OWNCD_FLAG == FALSE,
    MULTI_ADFORCD_FLAG == FALSE
  ),
  variables = c(
    "LAT",
    "LON",
    "DIA_est",
    "HT_est",
    "ACTUALHT_est",
    "YEAR",
    "OWNCD",
    "ADFORCD",
    "SPCD_CORR",
    "SPCD_FLAG",
    "STATUSCD"
  )
)

```

## Look for shrinking trees

```{r}

changes <- az_nfs |>
  group_by(TREE_COMPOSITE_ID) |>
  arrange(YEAR) |>
  mutate(LAST_YEAR = lag(YEAR),
         DIA_DIFF = DIA_est - lag(DIA_est),
         HT_DIFF = HT_est - lag(HT_est),
         AHT_DIFF = ACTUALHT_est - lag(ACTUALHT_est)) |>
  ungroup() |>
  
  arrange(TREE_COMPOSITE_ID, YEAR)

shrinking <- changes |> 
  filter(DIA_DIFF < 0 |
           HT_DIFF < 0 |
           AHT_DIFF < 0) 
```

```{r}
theme_set(theme_bw())
ggplot(shrinking, aes(DIA_DIFF)) +
  geom_histogram(bins = 50, boundary = 0) +
  geom_vline(xintercept = 0)

ggplot(shrinking, aes(HT_DIFF)) +
  geom_histogram(bins = 50, boundary = 0) +
  geom_vline(xintercept = 0)

ggplot(shrinking, aes(AHT_DIFF)) +
  geom_histogram(bins = 50, boundary = 0) +
  geom_vline(xintercept = 0)

```

```{r}

trees_that_ever_shrink <- shrinking$TREE_COMPOSITE_ID |> unique()

when_the_trees_shrank <- shrinking |>
  select(TREE_COMPOSITE_ID, INVYR) |> 
  distinct()

tree_records <- query_tree_surveys(con = con,
                                   conditions = create_conditions(TREE_COMPOSITE_ID %in% trees_that_ever_shrink),
                                   variables = c("DIA", "ACTUALHT", "HT", "STATUSCD", "ANY_SPCD_FLAG", 
                                                 "DAMAGE_AGENT_CD1",
                                                 "DAMAGE_AGENT_CD2",
                                                 "DAMAGE_AGENT_CD3",
                                                 "DSTRBCD1",
                "DSTRBCD2",
                "DSTRBCD3")) |>
  arrange(INVYR) |>
  mutate(DIA_DIFF = DIA - lag(DIA),
         HT_DIFF = HT - lag(HT),
         AHT_DIFF = ACTUALHT - lag(ACTUALHT),
         LAST_DIA = lag(DIA),
         LAST_HT = lag(HT),
         LAST_AHT = lag(ACTUALHT)) |>
  right_join(when_the_trees_shrank)

status_cd_when_shrank <- tree_records |>
  group_by(STATUSCD) |>
  tally()

status_cd_when_shrank

damage_dist_when_shrank <- tree_records |>
  mutate(across(contains("DAMAGE"), \(x) ifelse(is.na(x), 0, x))) |>
  mutate(across(contains("DAMAGE"), as.numeric)) |>
  mutate(HAS_DAMAGE = (DAMAGE_AGENT_CD1 + DAMAGE_AGENT_CD2 + DAMAGE_AGENT_CD3) > 0) |>
  mutate(across(contains("DSTRBCD"), \(x) ifelse(is.na(x), 0, x))) |>
  mutate(across(contains("DSTRBCD"), as.numeric)) |>
  mutate(IS_DISTURBED = (DSTRBCD1 + DSTRBCD2 + DSTRBCD3) > 0) |>
  mutate(dead_damaged_disturbed = ((STATUSCD == 2) + HAS_DAMAGE + IS_DISTURBED) > 0)

damage_dist_when_shrank |>
  group_by(HAS_DAMAGE, STATUSCD, IS_DISTURBED) |>
  tally()

damage_dist_when_shrank |> 
  group_by(dead_damaged_disturbed) |>
  tally()
```

3805 of 12820 incidents of shrinking coincide with being a dead tree. 

3524 of 12820 incidents coicide with at least one DAMAGE code. 

Another 727 are alive, undamaged, but marked as DISTURBED in some way. 

That stills leaves about 40% of instances where the tree is alive, not recorded as damaged, and shrinks on at least one measure (ht, actualht, dia).

```{r}

ggplot(damage_dist_when_shrank, aes(DIA_DIFF)) +
  geom_histogram(boundary = 0, bins = 50) +
  facet_wrap(vars(dead_damaged_disturbed), scales = "free_y", ncol = 1) +
  geom_vline(xintercept = 0)

ggplot(damage_dist_when_shrank, aes(HT_DIFF)) +
  geom_histogram(boundary = 0, bins = 50) +
  facet_wrap(vars(dead_damaged_disturbed), scales = "free_y", ncol = 1) +
  geom_vline(xintercept = 0)


ggplot(damage_dist_when_shrank, aes(AHT_DIFF)) +
  geom_histogram(boundary = 0, bins = 50) +
  facet_wrap(vars(dead_damaged_disturbed), scales = "free_y", ncol = 1) +
  geom_vline(xintercept = 0)


```

```{r}

mean(damage_dist_when_shrank$ANY_SPCD_FLAG)

```

About 5% of these trees have a SPCD_FLAG.

# Really big jumps

```{r}

tree_records_prop <- damage_dist_when_shrank |>
  mutate(DIA_PROP = DIA_DIFF / LAST_DIA,
         HT_PROP = HT_DIFF / LAST_HT,
         AHT_PROP = AHT_DIFF / LAST_AHT)

ggplot(tree_records_prop, aes(DIA_PROP)) +
  geom_histogram(boundary = 1, bins = 50) +
  facet_wrap(vars(dead_damaged_disturbed), scales = "free_y", ncol = 1) +
  geom_vline(xintercept = 1)

ggplot(tree_records_prop, aes(HT_PROP)) +
  geom_histogram(boundary = 1, bins = 50) +
  facet_wrap(vars(dead_damaged_disturbed), scales = "free_y", ncol = 1) +
  geom_vline(xintercept = 1)


ggplot(tree_records_prop, aes(AHT_PROP)) +
  geom_histogram(boundary = 1, bins = 50) +
  facet_wrap(vars(dead_damaged_disturbed), scales = "free_y", ncol = 1) +
  geom_vline(xintercept = 1)

```

```{r}


ggplot(damage_dist_when_shrank, aes(LAST_DIA, DIA)) +
  geom_point(alpha = .05) +
  geom_abline(slope = 1, intercept = 0) +
  facet_wrap(vars(dead_damaged_disturbed))

ggplot(damage_dist_when_shrank, aes(LAST_HT, HT)) +
  geom_point(alpha = .05) +
  geom_abline(slope = 1, intercept = 0) +
  facet_wrap(vars(dead_damaged_disturbed))

ggplot(damage_dist_when_shrank, aes(LAST_AHT, ACTUALHT)) +
  geom_point(alpha = .05) +
  geom_abline(slope = 1, intercept = 0) +
  facet_wrap(vars(dead_damaged_disturbed))


```

# Clean up

```{r}
dbDisconnect(con, shutdown = TRUE)

```
