---
title: "Annualized estimates taking into account mortality"
format: gfm
editor: source
---

```{r, output = F, message = F, warning = F}

source(here::here("R", "query_tables_db_fxns.R"))
source(here::here("R", "query_annualized.R"))

con <- connect_to_tables(here::here("data", "db", "foresttime-to-share.duckdb"))

```

```{r, output = F}

az_nfs_annualized <- query_annualized(con,
                                      conditions = create_conditions(
                                        STATECD == 4,
                                        OWNCD == 11,
                                        MULTI_OWNCD_FLAG == FALSE,
                                        MULTI_ADFORCD_FLAG == FALSE
                                      ),
                                      variables = c("DIA_est", "HT_est", "ACTUALHT_est", "DIA_est_mort", "HT_est_mort", "ACTUALHT_est_mort", "YEAR", "ADFORCD", "SPCD_CORR", "SPCD_FLAG", "STATUSCD", "DIES"))

```

This plot illustrates annual estimates of tree diameter taking into account mortality. 
For trees that die and have a MORTYR recorded, I recalculated the interpolated diameter estimates assuming they stopped changing in diameter in MORTYR.
This is shown in the dotted lines on the graph and how they diverge from the solid lines.
Note that some trees die and do not have MORTYR recorded, or have a MORTYR that is the same as the last INVYR. 
For these trees, the diameter estimates do not change whether or not you take into account MORTYR.

```{r, echo = F}
library(ggplot2)

ggplot(filter(az_nfs_annualized, PLOT_COMPOSITE_ID == "4_1_11_80552"), aes(YEAR, DIA_est_mort, group = TREE_COMPOSITE_ID, color = DIES)) +
  geom_line(linetype = 2) +
  geom_line(aes(y = DIA_est)) +
  geom_point() + 
  theme_bw() +
  scale_color_viridis_d(end = .8)


```


```{r}

dbDisconnect(con, shutdown = TRUE)

```


