---
title: "Extracting FIA timeseries"
format: gfm
editor: source
---

```{r}

library(duckdb)
library(dplyr)
library(ggplot2)

source(here::here("R", "query_tables_db_fxns.R"))

theme_set(theme_bw())

```


# Whitebark pine for Montana, Idaho, and Wyoming

SPCD = 101

STATECD = 16, 30, 56

```{r}

con <- connect_to_tables(here::here("data", "db", "forestTIME-whitebark.duckdb"))

whitebark_pine <-
  query_tables_db(con = con,
  conditions = create_conditions(STATECD %in% c(16, 30, 56),
                    SPCD == 101,
                    SPCDS == 1),
  variables = c("STATUSCD", "DIA", "HT", "COND_STATUS_CD", "LAT", "LON",
                "BALIVE", "SICOND", "SISP", "SIBASE",
                "DSTRBCD1",
                "DSTRBYR1",
                "DSTRBCD2",
                "DSTRBYR2",
                "DSTRBCD3",
                "DSTRBYR3",
                "SDIMAX_RMRS",
                "SDI_RMRS",
                "SLOPE",
                "ASPECT",
                "CONDPROP_UNADJ",
                "RECONCILECD")
)

dbDisconnect(con, shutdown = TRUE)

knitr::kable(head(whitebark_pine))

```


## How many trees have been surveyed how many times in each state?

```{r}
#| echo: FALSE
#| message: FALSE
#| warning: FALSE

whitebark_pine_measures <- whitebark_pine |> 
  group_by(STATECD, TREE_UNIQUE_ID) |>
  summarize(n = length(unique(INVYR))) 

knitr::kable(whitebark_pine_measures |>
               group_by(STATECD, n) |>
               tally() |>
               rename(n_measures = n,
                      n_trees = nn))


```

**Notably, zero trees in Wyoming (STATECD = 56) have been surveyed more than one time.**

## Plotting DIA, HT of trees with repeated measurements

```{r}
#| echo: FALSE
#| message: FALSE
#| warning: FALSE

whitebark_pine_repeats <- filter(whitebark_pine_measures, n > 1) |>
  left_join(whitebark_pine)

ggplot(whitebark_pine_repeats, aes(INVYR, DIA, group = TREE_UNIQUE_ID)) +
  geom_line() +
  facet_wrap(vars(STATECD), scales = "free") +
  ggtitle("DIA")


ggplot(whitebark_pine_repeats, aes(INVYR, HT, group = TREE_UNIQUE_ID)) +
  geom_line() +
  facet_wrap(vars(STATECD), scales = "free") +
  ggtitle("HT")

```

## Saving data to share

```{r}

write.csv(whitebark_pine, here::here("use_cases", "whitebark_pine", "whitebark_pine.csv"))

```

The saved file is `r trunc(file.size(here::here("use_cases", "whitebark_pine", "whitebark_pine.csv")) / 1E6, 3)` MB.